---
title: "Abstract classification in Macroecology"
author: "Alexander Florez Rodriguez"
date: "11/22/2018"
output: md_document
---

```{r echo=F, message=F}
library(keras)
library(knitr)
# Chunk to classify abstracts by journal
read_chunk("class_abstract_binary.R")
```

## The stupid question

If I give you an abstract to read (just the text, forget about the format), would 
you be able to tell me in what journal was published?

Let's start easy. You only have 2 options (50% chance): you have to differentiate
between abstracts in "Diversity and Distributions" and "Journal of Biogeography". 
Two staple journals in Macroecology!

## The stupid idea

Here I use neural networks to classify abstracts by journal. I used all the abstracts
for "Diversity and Distributions" and "Journal of Biogeography"
available in ISI Web of knowledge to train, test and validate the neural network.

## Can you guess the accuracy of the neural network?

\vspace{12pt}
\vspace{12pt}
```{r get.abstracts, echo=F}

```

```{r data.segmentation, echo=F}

```

```{r tokenization, echo=F}

```

### Neural network model using a embedding layer
```{r NN.define.model}

```

```{r NN.compile.model}

```

```{r plot.validation, fig.height=6, fig.width=12}
```
\vspace{12pt}
The maximum accuracy of this model is `r max(history$metrics$val_acc)`

\vspace{12pt}
\vspace{12pt}
### Neural network model using a embedding layer + a densely connected layer
```{r adding.dense.layer, fig.height=6, fig.width=12}
```
\vspace{12pt}
The maximum accuracy of this model is `r max(history$metrics$val_acc)`

### Neural network model using a embedding layer + a densely connected layer + a longer text sequence
```{r longer.sequence.lenght, fig.height=6, fig.width=12}
```
The maximum accuracy of this model is `r max(history$metrics$val_acc)`

## The stupid result
Using these simple neural networks we can classify abstracts by journal with around 83% of accuracy...
Now, can we use the same approach to classify papers by topic (e.g. latitudinal gradient)? or by traits? or
by taxonomic group?
